\documentclass[10pt,a4paper]{article}
\usepackage{fullpage}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage[shortlabels]{enumitem}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{ulem}
\usepackage{tikz}
\usetikzlibrary{positioning}
\newcommand{\nl}{\vspace{12pt}}
\usepackage[utf8]{inputenc}
\usepackage{float}
%\usepackage{minted}
% Catpuccin Macchiato 

\newcommand{\M}[1]{\mathbf{#1}}
\newcommand{\V}[1]{\mathbf{#1}}
\newcommand{\Vh}[1]{\mathbf{\hat{#1}}}
\newcommand{\norm}[1]{\left | \left | #1 \right | \right |}
\newcommand{\RR}{\mathbbm{R}}        % set of real numbers

\setlength{\parindent}{0pt}


\title {Drone}
\author {Ken Schlesiger}
\begin{document}
\maketitle
\newpage
\section*{Hardware}
\label{sec:hardware}

\subsection*{Structure}
\label{structure}
The basic structure will be 3D printed. % TODO everything

\subsection*{Flight-controller}
The flight-controller is based on the STM32f407G. 

\subsection*{Sensors}
Currently the intend is to use SPI to communicate with all Sensors to ensure fast and reliable communication.

\subsubsection*{Attitude Sensors}
The Drone will use an LSM9DS1 containing an Accelerometer, Gyroscope and Magnetometer for attitude determination.

\subsubsection*{Position Sensors}

\subsection*{Power}
\label{sec:power}

Ill need a charger, perhaps
%https://www.amazon.de/SKYRC-B6DC-Ladeger%C3%A4t-LiFePO-LiIon/dp/B0822RM9VJ?source=ps-sl-shoppingads-lpcontext&ref_=fplfs&psc=1&smid=A3JWKAKR8XB7XF&language=de_DE&gQT=2

\subsubsection*{Batteries}

\subsubsection*{Power Distribution and Regulation}

\subsubsection{Power conditioning}

\subsection*{Motors}
\label{motors}

Currently unclear, probably something like iFlight XING2 2306 2555KV 4S FPV Motor. 
d.h. 4s lipo



\section*{Software}
\label{sec:software}
The SPI bus should mainly be concerned with handling Attitude relevant sensor information and sending control commands.
All other sensors should be only queried when necessary.

\subsection*{Flight-controller}
The flight-controller is based on an STM32f407G ruining \href{https://github.com/kenscl/SROS}{SROS}.
Currently the intend is to communicate through all external devices using SPI.

\subsubsection*{Attitude Determination}
Using the LSM9DS1 Gyroscope and Accelerometer data is gathered every 3 ms, while magnetometer data is gathered when it becomes available (about every 50ms).
An EKF % Link to document
is used to determine the Attitude. 

\end{document}
